# **ANOMALY DETECTION**

## **Objective**

To check whether there are any anomalies in the given sales dataset.

## **Context**

The dataset contains sales data collected over a period of four months in 2019.The task is to detect if there are any anomalies in the sales totals

## **Loading libraries**

```{r}
library(tidyverse)
library(anomalize)
```

##**Loading and Previewing data**
```{r}
sales_df <- read.csv("http://bit.ly/CarreFourSalesDataset")
```
**First 6 records**
```{r}
rmarkdown::paged_table(head(sales_df, n=5))
```

**Last 6 records** 
```{r}
rmarkdown::paged_table(tail(sales_df, n=5))
```

**Dataset Dimension**
```{r}
#The data has 1000 records and 2 features
dim(sales_df)
```

**Data types**
```{r}
sapply(sales_df,class)
```
```{r}
sales_df$Date <- as.Date(sales_df$Date, "%m/%d/%y")
sapply(sales_df,class)
```

**Column names**
```{r}
colnames(sales_df)
```
## **Data Cleaning**
**Duplicates**
```{r}
#Checking for duplicated records
sum(duplicated(sales_df))
#The dataset has no duplicates
```
**Missing Values**
```{r}
#Checking for missing values
colSums(is.na(sales_df))
#The dataset has no missing values
```
**Outliers**
```{r}
#Checking for outliers
boxplot(sales_df$Sales)
```
**Distribution**
```{r}
#Checking the distribution of the sales column
qqnorm(sales_df$Sales,main = "Sales Distribution")
qqline(sales_df$Sales, lwd=2.5,col="red")
#The sales column datapoint are normaly distributed
```
```{r}
sales_df <- sales_df %>% 
  group_by(Date) %>% 
  tally()
colnames(sales_df) <- c('Date','Total_Sales')
head(sales_df)
```
## **Detecting Anomalies**
```{r}
anomalized<-sales_df %>%
    time_decompose(Total_Sales) %>%
    anomalize(remainder) %>%
    time_recompose() %>%
    plot_anomalies(ncol = 3, alpha_dots = 2.5)
anomalized
```


```{r}
anomalized<-sales_df %>%
    time_decompose(Total_Sales, merge = TRUE) %>%
    anomalize(remainder) %>%
    time_recompose() %>%
    plot_anomaly_decomposition() +
  ggtitle("Anomaly Distribution")
anomalized
```
```

**Summary:**

There are no anomalies detected
