# **ASSOCIATION ANALYSIS**

## **Objective**

To create association rules that will allow you to identify relationships between variables in the dataset.

## **Context**

The dataset contains data of items bought by different customers. The project aims to create a model using association rule to predictitem purchasing behaviors

## **Loading libraries**
```{r}
library(arules)
```
##**Loading and Previewing data**
```{r}
url <-"http://bit.ly/SupermarketDatasetII"
order<-read.transactions(url, sep = ",")
```
**First 6 records**
```{r}
orders<-as.data.frame(itemLabels(order))
colnames(orders) <- "Item"
head(orders, 6)
```
**First 6 orders**
```{r}
inspect(order[1:6])
```

**Last 6 records** 
```{r}
orders<-as.data.frame(itemLabels(order))
colnames(orders) <- "Item"
tail(orders, 6)
```

**Dataset Dimension**
```{r}
dim(order)
```

**Data types**
```{r}
class(order)
```

## **Data Cleaning**
**Duplicates**
```{r}
#Checking for duplicated records
sum(duplicated(order))
#dropping duplicates
order<-order[!duplicated(order)]
order
```
## **Exploratory Data Analysis**
```{r}
summary(order)
#Most frequently purchased items are mineral water, spaghetti,eggs,chocolate and french fries
```
```{r}
#Displaying the top ten most purchased items.
itemFrequencyPlot(order, type = "absolute",topN=10,col=rainbow(10),main = "Most Purchased Items")
```
## **Building Model**
```{r}
#Applying the arules algorithim with 80% confidence level
model<- apriori (order, parameter = list(supp = 0.001, conf = 0.8))
#A model with 247 rules is created
model
```
```{r}
#Describing the model
summary(model)
#The model has a minimum of three items and a maximum of six items
```
```{r}
#Showing the first five record and the predicted items likely to be bought
inspect(model[1:5])
# A person who buys bugers,hand protein bar has the likelihood of 85.7% of buying mineral water
```

```{r}
#sorting by confidence level
model<-sort(model, by="confidence", decreasing=TRUE)
inspect(model[1:5])
#A person buying nonfat milk and soup is 100% likely to buy mineral water

```
```{r}
#Predicting items likely to be bought buy a customer who has pick mineral water as an item
`mineral water` <- subset(model, subset = lhs %pin% "mineral water")

# Order by confidence
`mineral water`<-sort(`mineral water`, by="confidence", decreasing=TRUE)

# inspect top 10
inspect(`mineral water`[1:10])
```

